Argennon: A Scalable Smart Contract Platform

### 2022-03-31 tex/sections/avm.tex
Remove usages of "ZK-EDB"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

The Argennon\footnote{The classical pronunciation should be used:\textipa{/Ar"gen.non/}} Smart Contract Execution
Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ applications).
An Argennon application essentially is a http server whose state is kept in the Argennon blockchain and
its logic is described using an Argennon Standard Representation (ASR).

An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
for the architecture and requirements of the Argennon platform.
Argennon supports two standard representations: one is a high level text based language which needs
costly compilation before being executed on a hardware machine. The other is a low level binary representation which
usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
language is intended for preserving high level information of applications logic to facilitate
platform specific compiler optimization at host nodes. On the other hand, the low level language is designed for
efficient execution of applications that are not frequently used.

The state of an Argennon application is stored in byte addressable finite arrays of memory called
\emph{heap chunks}. An application can have several heap chunks with different sizes, and can remove or
resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
by its owner. In addition to heap chunks, every application also has an amount of non-persistent local memory for
storing temporary data.

@@

The Argennon Smart Contract Execution Environment can be seen as a machine for executing http requests, producing
their http response and updating related heap chunks. The AscEE executes requests sequentially \footnote{Actually
requests are executed in parallel but by performing data dependency analysis the result is guaranteed to be identical
with sequential execution of requests.} and each request has a separate \emph{execution session}. Execution sessions
are separate sessions of executing smart contract's code in order to fulfill \emph{external} http requests. An
external request is a request that is not made by another Argennon application.

@@

During an execution session an application can make \emph{internal} http requests to other applications. Those
requests will not start a new execution session and will be executed within the current session. In AscEE making a
http request to an application is similar to a function invocation. For that reason, we also refer to internal requests
as application calls.

@@

Attachment of a request is a list of request identifiers of the current block that are "attached" to the request.
That means, for validating that request a validator first needs to "inject" the digest of attached requests into the
http request.

@@

The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
session is always measured per session and default pre-defined resource caps are used. On the other hand in a monitored
session, some resources are measured per application call and some
caps are determined dynamically by the external http request (i.e.\ transaction). An optimistic session must always
complete normally, a monitored session is allowed to either complete normally or abruptly.

@@

    \item \textbf{local memory}:
    any memory usage of an application that is not part of a heap chunk and is not part of another resource
    will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
    request and returns the http response (i.e.\ the application call completes) its local memory is deleted.
    This resource is measured in bytes.

>>>>

The Argennon\footnote{The classical pronunciation should be used:\textipa{/Ar"gen.non/}} Smart Contract Execution
Environment (AscEE) is an abstract execution environment for executing Argennon smarts contracts (a.k.a\ applications).
An Argennon application essentially is an HTTP server whose state is kept in the Argennon blockchain and
its logic is described using an Argennon Standard Representation (ASR).

An Argennon Standard Representation (ASR) is a programming language for describing argennon applications, optimized
for the architecture and requirements of the Argennon platform.
Argennon supports two standard representations: one is a high level text based language which needs
costly compilation before being executed on a hardware machine. The other is a low level binary representation which
usually can be executed, with minimal pre-processing, by a JIT compiler or an emulator. The high level
language is intended for preserving the high level information of applications logic to facilitate
platform specific compiler optimization at host nodes. On the other hand, the low level language is designed for
efficient execution of applications that are not frequently used.

The state of an Argennon application is stored in byte addressable finite arrays of memory called
\emph{heap chunks}. An application may have several heap chunks with different sizes, and can remove or
resize its heap chunks or allocate new chunks. Every chunk belongs to exactly one application and can only be modified
by its owner. In addition to heap chunks, every application also has an amount of non-persistent local memory for
storing temporary data.

@@

The Argennon Smart Contract Execution Environment can be seen as a machine for executing HTTP requests, producing
their HTTP response and updating related heap chunks. The AscEE executes requests sequentially \footnote{Actually
requests are executed in parallel but by performing data dependency analysis the result is guaranteed to be identical
with sequential execution of requests.} and each request has a separate \emph{execution session}. Execution sessions
are separate sessions of executing smart contract's code in order to fulfill \emph{external} HTTP requests. An
external request is a request that is not made by another Argennon application.

@@

During an execution session an application can make \emph{internal} HTTP requests to other applications. Those
requests will not start a new execution session and will be executed within the current session. In AscEE making a
HTTP request to an application is similar to a function invocation. For that reason, we also refer to internal requests
as application calls.

@@

Attachment of a request is a list of request identifiers of the current block that are "attached" to the request.
That means, for validating that request a validator first needs to "inject" the digest of attached requests into the
HTTP request.

@@

The AscEE has two type of execution sessions: \emph{optimistic} and \emph{monitored}. Resource usage of an optimistic
session is always measured per session and default pre-defined resource caps are used. On the other hand in a monitored
session, some resources are measured per application call and some
caps are determined dynamically by the external HTTP request (i.e.\ transaction). An optimistic session must always
complete normally, a monitored session is allowed to either complete normally or abruptly.

@@

    \item \textbf{local memory}:
    any memory usage of an application that is not part of a heap chunk and is not part of another resource
    will be considered as local memory usage. Local memory is not persistent and when an application finishes serving a
    request and returns the HTTP response (i.e.\ the application call completes) its local memory is deleted.
    This resource is measured in bytes.

### 2022-03-31 tex/sections/persistence.tex
Remove usages of "ZK-EDB"
Signed-off-by: aybehrouz <behrouz_ayati@yahoo.com>

The Argennon Virtual Machine has two persistent memory areas: \emph{method area}, and \emph{heap}. Method area stores
bytecodes of methods\footnote{also it stores constant area blocks.}, and heap stores memory chunks. Both of these
data elements, bytecodes and chunks, can be considered as continuous pieces of byte addressable memory. Throughout this
chapter, we shall call these data elements \emph{objects}.

@@

In the AVM persistence layer, similar objects are clustered together and constitute a bigger data element which we call
a \emph{page}.\footnote{we avoid calling them clusters, because usually a cluster refers to a \emph{set}. AVM object
clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
A page is an ordered list of an arbitrary number of objects, which their order reflects the order they were added to
the page:
\[
    P = (O_1,O_2,\dots,O_n),\quad i < j \; \Leftrightarrow \; \textrm{$O_i$ was added before $O_j$}\ .
\]

A page of the AVM storage should contain objects that have very similar access pattern. Ideally, when a page is needed
for validating a block, almost all of its objects should be needed for either reading or writing. We prefer
that the objects are needed for the same access type. In other words, the objects of a page are chosen in a way that
for validating a block, we usually need to either read all of them or modify\footnote{and probably read.} all of them.

\section{Zero-knowledge Databases}\label{sec:zk-edb}

Pages of the AVM storage are persisted using updatable zero-knowledge elementary databases (ZK-EDB). Argennon
has three zero-knowledge databases: \emph{staking} database, which stores all the data that is associated with
the Argennon consensus protocol. \emph{method} database, which stores the AVM method area, and \emph{heap} database,
which stores the AVM heap. The commitment of these three ZK-EDBs are included in every block of the Argennon blockchain.

We consider the following properties for a ZK-EDB:
\begin{itemize}
    \item The ZK-EDB contains a mapping from a set of keys to a set of values.
    \item Every state of the database has a commitment \(C\).
    \item The ZK-EDB has a method \((D, \pi) = \text{get}(x)\), where \(x\) is a key and \(D\) is the associated data
    with \(x\), and \(\pi\) is a proof.
    \item A user having \(C\) and \(\pi\) can verify that \(D\) is really associated with \(x\), and \(D\) is not
    altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the ZK-EDB\@.
    \item Having \(\pi\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
    associated with \(x\) instead of \(D\).
\end{itemize}

Pages of the AVM storage are stored in the ZK-EDBs, with an index: \texttt{pageIndex} as their key.
The \texttt{pageIndex} is required to be smaller than a certain value, determined by the
protocol, to facilitate the usage of ZK-EDBs that are based on vector commitments.
For this reason, the AVM clustering algorithm always tries to reuse indices and keep the number of used indices
as low as possible.

The commitments of the AVM ZK-EDBs are affected by the way data objects are clustered. Therefore,
the Argennon clustering algorithm has to be a part of the consensus protocol.

@@

A block proposer could obtain clustering directives from any third party source\footnote{we can say the AVM clustering
algorithm is essentially off-chain.}. This will not
affect Argennon security, since the integrity of a database can not be altered by clustering directives.
Those directives can only affect the performance of the Argennon network, and directives of a single block can
not affect the performance considerably.

>>>>

The Argennon Smart Contract Execution Environment has two persistent memory areas: \emph{code area}, and \emph{heap}.
Code area stores the Argennon Standard Representation of
applications\footnote{also it stores applications' constants.}, and heap stores heap chunks.
Both of these data elements, ASRs and chunks, can be considered as continuous arrays of bytes.
Throughout this chapter, we shall call these data elements \emph{objects}.

@@

In the AscEE persistence layer, similar objects are clustered together and constitute a bigger data element which we
call
a \emph{page}.\footnote{we avoid calling them clusters, because usually a cluster refers to a \emph{set}. AscEE object
clusters are not sets. They are ordered lists, like a page containing an ordered list of words or sentences.}
A page is an ordered list of an arbitrary number of objects, which their order reflects the order they were added to
the page:
\[
    P = (O_1,O_2,\dots,O_n),\quad i < j \; \Leftrightarrow \; \textrm{$O_i$ was added before $O_j$}\ \cdot
\]

A page of the AscEE storage should contain objects that have very similar access pattern. Ideally, when a page is needed
for validating a block, almost all of its objects should be needed for either reading or writing. We prefer
that the objects are needed for the same access type. In other words, the objects of a page are chosen in a way that
for validating a block, we usually need to either read all of them or modify\footnote{and probably read.} all of them.

\section{Publicly Verifiable Database Servers}\label{sec:zk-edb}

Pages of the AscEE storage are persisted using \emph{dynamic cryptographic accumulators}. Argennon
has three dynamic cryptographic accumulators: \emph{staking} database, which stores all the data that is associated with
the Argennon consensus protocol. \emph{code} database, which stores the AscEE code area, and \emph{heap} database,
which stores the AscEE heap. The commitment of these three accumulators are included in every block of the Argennon
blockchain. These accumulators, in the Argennon network, are hosted on special database nodes called Publicly
Verifiable Data Base (PV-DB) servers.
We consider the following properties for a PV-DB:
\begin{itemize}
    \item The PV-DB contains a mapping from a set of keys to a set of values.
    \item Every state of the database has a commitment \(C\).
    \item The PV-DB has a method \((D, \pi) = \text{get}(x)\), where \(x\) is a key and \(D\) is the associated data
    with \(x\), and \(\pi\) is a proof.
    \item A user having \(C\) and \(\pi\) can verify that \(D\) is really associated with \(x\), and \(D\) is not
    altered. Consequently, a user who can obtain \(C\) from a trusted source does not need to trust the PV-DB\@.
    \item Having \(\pi\) and \(C\) a user can compute the commitment \(C'\) for the database in which \(D'\) is
    associated with \(x\) instead of \(D\).
\end{itemize}

Pages of the AscEE storage are stored in the PV-DBs, with an index: \texttt{pageIndex} as their key.
The \texttt{pageIndex} is required to be smaller than a certain value, determined by the
protocol, to facilitate the usage of PV-DBs that are based on vector commitments.
For this reason, the AscEE clustering algorithm always tries to reuse indices and keep the number of used indices
as low as possible.

The commitments of the AscEE cryptographic accumulators are affected by the way data objects are clustered. Therefore,
the Argennon clustering algorithm has to be a part of the consensus protocol.

@@

A block proposer could obtain clustering directives from any third party source\footnote{we can say the AscEE clustering
algorithm is essentially off-chain.}. This will not
affect Argennon security, since the integrity of a database can not be altered by clustering directives.
Those directives can only affect the performance of the Argennon network, and directives of a single block can
not affect the performance considerably.
